---
title: "Introduction to stJoincount"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to stJoincount}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r set up, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message = FALSE, warning=FALSE}
library(stJoincount)
library(spdep)
library(raster)
library(pheatmap)
library(ggplot2)
```

v0.99.6

stJoincount: Quantification tool for spatial correlation between clusters in spatial transciptomics pre-processed data using join count statustic approach.  

# Introducation
Spatial dependency is the relationship between locational and attribute similarity. The measure reflects whether an attribute of a variable observed at one location is independent of values observed at neighboring locations. Positive spatial dependency exists when neighboring attributes are more similar than what could be explained by chance. Likewise, a negative spatial dependency is reflected by a dissimilarity of neighboring attributes. Join count analysis allows for quantification of the spatial dependencies of nominal data in an arrangement of spatially adjacent polygons.

This tool requires data produced with the 10X Genomics Visium Spatial Gene Expression platform with customized clusters. The purpose of this R package is to perform join count analysis for spatial correlation between clusters. Examples of how to run this tool are below:  

# Loading the data
In this vignette, we are going to use an human breast cancer spatial transcriptomics sample.
```{r}
fpath <- system.file("script", "humanBC.rda", package="stJoincount")
load(fpath)
head(humanBC)
```

This tool first converts a labeled spatial tissue map into a raster object, in which each spatial feature is represented by a pixel coded by label assignment. This process includes automatic calculation of optimal raster resolution and extent for the sample.
```{r}
resolutionList <- resolutionCalc(humanBC)
resolutionList
```

```{r}
mosaicIntegration <- rasterizeEachCluster(humanBC)
```

# Visualization
After the labeled spatial sample being converted, the raster map can be visualized by:
```{r, fig.dim = c(6, 6)}
mosaicIntPlot(humanBC, mosaicIntegration)
```

# Join count analysis
A neighbors list is then created from the rasterized sample, in which adjacent and diagonal neighbors for each pixel are identified. After adding binary spatial weights to the neighbors list, a multi-categorical join count analysis is performed to tabulate "joins" between all possible combinations of label pairs. The function returns the observed join counts, the expected count under conditions of spatial randomness, and the variance calculated under non-free sampling.
```{r}
joincount.result <- joincountAnalysis(mosaicIntegration)
```

The z-score is then calculated as the difference between observed and expected counts, divided by the square root of the variance. A heatmap of z-scores represents the result from the join count analysis for all possible label pairs.

```{r, fig.dim = c(6, 6)}
matrix <- zscoreMatrix(humanBC, joincount.result)
zscorePlot(matrix)
```

```{r}
sessionInfo()
```
